Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2025-01-14 paper/main.tex
cleaned up phrasing of fig 6 within-lecture results

We observed a similar set of results when we restricted our estimates of
participants' knowledge for questions about each lecture to consider only
their performance on other questions about the \textit{same} lecture.
Specifically, for Quiz~1, participants' knowledge of \textit{Four Fundamental
Forces}-related questions, estimated from their performance on other
\textit{Four Fundamental Forces}-related questions, was predictive of their
ability to answer those questions correctly ($OR = 15.934,\ 95\%\
\textnormal{CI} = [5.173,\ 38.005],\ \lambda_{LR} = 40.971,\ p = 0.001$). The
same was true of participants' estimated knowledge for \textit{Birth of
Stars}-related questions based on their performance on other \textit{Birth of
Stars}-related questions ($OR = 9.775,\ 95\%\ \textnormal{CI} = [2.93,\
25.08],\ \lambda_{LR} = 13.924,\ p = 0.001$). These within-lecture knowledge
estimates also predicted success on questions about both lectures when we
computed them analogously for Quiz~2 (\textit{Four Fundamental Forces}: $OR =
35.126,\ 95\%\ \textnormal{CI} = [5.113,\ 123.868],\ \lambda_{LR} = 32.251,\
p < 0.001$; \textit{Birth of Stars}: $OR = 4.717,\ 95\%\ \textnormal{CI} =
[2.021,\ 9.844],\ \lambda_{LR} = 16.788,\ p < 0.001$). For Quiz~3, we found that
within-lecture knowledge estimates predicted participants' success on
\textit{Birth of Stars}-related questions ($OR = 16.902,\ 95\%\ \textnormal{CI}
= [3.353,\ 53.265],\ \lambda_{LR} = 23.233,\ p < 0.001$) but not on \textit{Four
Fundamental Forces}-related questions ($OR = 2.485,\ 95\%\ \textnormal{CI} =
[0.724,\ 8.366],\ \lambda_{LR} = 1.984,\ p = 0.170$). This may indicate that the
within-lecture knowledge estimates are susceptible to ceiling effects in
participants' quiz performance. On Quiz~3, after viewing both lectures, no
participant answered more than three \textit{Four Fundamental Forces}-related
questions incorrectly, and all but five participants (out of 50) answered two or
fewer questions incorrectly. (This was the only subset of questions about either
lecture, across all three quizzes, for which this was true.) Because of this,
when we held out one incorrectly answered
\textit{Four Fundamental Forces}-related question from a given participant's
Quiz~3 responses and estimated their knowledge at its embedding coordinate using
the remaining \textit{Four Fundamental Forces}-related questions they answered,
for 90\% of participants, that estimate leveraged information about at most a
single other question they were \textit{not} able to correctly answer. This
homogeneity in participants' success on questions used to estimate their
knowledge may have hurt our ability to accurately characterize the specific (and
by Quiz~3, relatively few) aspects of the lecture content they did \textit{not}
know about. Taken together, these results suggest that our knowledge estimates
can reliably distinguish between questions about different content covered by a
single lecture, provided there is sufficient diversity in participants' quiz
responses to extract meaningful information about both what they know and what
they do not know.

>>>>

We observed a similar set of results when we restricted our estimates of
participants' knowledge for questions about each lecture to consider only
their performance on other questions about the \textit{same} lecture.
Specifically, for Quiz~1, participants' knowledge of \textit{Four Fundamental
Forces}-related questions, estimated from their performance on other
\textit{Four Fundamental Forces}-related questions, was predictive of their
ability to answer those questions correctly ($OR = 15.934,\ 95\%\
\textnormal{CI} = [5.173,\ 38.005],\ \lambda_{LR} = 40.971,\ p = 0.001$). The
same was true of participants' estimated knowledge for \textit{Birth of
Stars}-related questions based on their performance on other \textit{Birth of
Stars}-related questions ($OR = 9.775,\ 95\%\ \textnormal{CI} = [2.93,\
25.08],\ \lambda_{LR} = 13.924,\ p = 0.001$). These within-lecture knowledge
estimates also predicted success on questions about both lectures when we
computed them analogously for Quiz~2 (\textit{Four Fundamental Forces}: $OR =
35.126,\ 95\%\ \textnormal{CI} = [5.113,\ 123.868],\ \lambda_{LR} = 32.251,\
p < 0.001$; \textit{Birth of Stars}: $OR = 4.717,\ 95\%\ \textnormal{CI} =
[2.021,\ 9.844],\ \lambda_{LR} = 16.788,\ p < 0.001$). For Quiz~3, we found that
within-lecture knowledge estimates predicted participants' success on
\textit{Birth of Stars}-related questions ($OR = 16.902,\ 95\%\ \textnormal{CI}
= [3.353,\ 53.265],\ \lambda_{LR} = 23.233,\ p < 0.001$) but not on \textit{Four
Fundamental Forces}-related questions ($OR = 2.485,\ 95\%\ \textnormal{CI} =
[0.724,\ 8.366],\ \lambda_{LR} = 1.984,\ p = 0.170$). This may indicate that the
within-lecture knowledge estimates are susceptible to ceiling effects in
participants' quiz performance. On Quiz~3, after viewing both lectures, no
participant answered more than three \textit{Four Fundamental Forces}-related
questions incorrectly, and all but five participants (out of 50) answered two or
fewer incorrectly. (This was the only subset of questions about either lecture, 
across all three quizzes, for which this was true.) Because of this, for 90\% of 
participants, our within-lecture estimates of their knowledge for \textit{Four 
Fundamental Forces}-related questions that they answered incorrectly leveraged 
information from at most a single other question they were \textit{not} able to 
correctly answer. This likely hampered our ability to accurately characterize 
the specific (and by the time they took Quiz~3, relatively few) aspects of the 
lecture content these participants did \textit{not} know about, and successfully 
distinguish them from the far more numerous aspects of the lecture content they 
now \textit{did} know about. Taken together, these results suggest that our 
knowledge estimates can reliably distinguish between questions about different
content covered by a single lecture, provided there is sufficient diversity in 
participants' quiz responses to extract meaningful information about both what 
they know and what they do not know.
